Gradual Soundness: Lessons from Static Python

### 2021-10-19 ecoop-2022/paper.tex
ecoop: contributions pass

In short, academic researchers are working to close the performance gap
without overly restricting the promise of gradual typing.
Industry researchers are, for the most part, sidestepping the problem with unsound types.

@@

\begin{itemize}
  \item
    Evidence that gradual soundness can work.
    \SP{} performs well on a microbenchmark suite and, more importantly, in production.
    The effort needed to convert code to \SP{} is low thanks to its novel application
    of concrete types.
  \item
    A formal description and analysis of the \SP{} core language.
    The analysis includes a proof of soundness and (perhaps) a metric for
    the relative cost of a type migration.
  \item
    A full list of the limitations in \SP{} relative to idealized gradual type
    systems and an informal description of the programming discipline that Instagram
    follows in their gradually sound codebase.
\end{itemize}

@@

We have written this paper to promote two future directions.
First, we want to encourage system-builders to reproduce the essentials of
\SP{} in another context and measure the net performance.
Second, we want to lend focus to theoretical research on gradual typing.
The restrictions that \SP{} adopts (relative to the ideal promise of gradual typing)
may be worthwhile assumptions for theory going forward.
On the flip side, there are issues in today's \SP{} that might
be addressed by future research.

>>>>

In short, academic researchers are working to close the performance gap
without overly restricting the promise of gradual typing.
Industry researchers are sidestepping the problem with unsound types---for the most part.

@@

\begin{itemize}
  \item
    Evidence that gradual soundness delivers fast performance incrementally,
    as programmers migrate to sound types.
    \SP{} can be applied to a codebase with no refactoring and no performance overhead.
    After refactoring to sound types, \SP{} performs well on a microbenchmark
    suite and in production.
  \item
    A formal description and analysis of the \SP{} core language.
    The analysis includes a proof of type soundness,
    a proof that optimization soundly improves code,
    and [FILL perhaps] a metric to quantify the refactoring cost and optimization benefit
    of a type migration.
  \item
    A list of the expressiveness limitations in \SP{} relative to
    idealized gradual type systems,
    and an informal description of the programming discipline that Instagram
    follows in their gradually-sound codebase.
\end{itemize}

@@

We have written this paper with two audiences in mind.
First, we want to encourage system-builders to reproduce the compelling
\SP{} language design in another context and measure performance.
Second, we want to lend focus to theoretical research on gradual typing.
The restrictions that \SP{} adopts (relative to the ideal promise of gradual typing)
may be worthwhile assumptions for theory going forward.
On the flip side, there are issues in today's \SP{} that might
be addressed by future research.
